---
title: "NBA Salary Correlation"
author: "Mohamed Karboj & Fouad Abdulghani"
format: html
editor: visual
---

```{r}
library(tidyverse)

nba <- read.csv("nba_2022-23_all_stats_with_salary.csv")
```

# Research Questions:

Do players who score more points per game earn higher salaries?\
Is there a significant difference in average salary between point guards or shooting guards?

## Descriptive Analysis

#### Histogram

```{r}
nba |>
  ggplot(aes(x = PTS)) +
  geom_histogram(fill = "steelblue", color = "white", bins = 30) +
  labs(title = "Distribution of Points per Game", x = "Points per Game", y = "Frequency")
```

#### Box Plot

```{r}
nba |>
  ggplot(aes(x = Position, y = Salary, fill = Position)) +
  geom_boxplot() +
  labs(title = "Salary Distribution by Position", x = "Position", y = "Salary") +
  theme_minimal()
```

#### Scatter Plot

```{r}
nba |> 
  ggplot(aes(x = PTS, y = Salary)) +
  geom_point(alpha = 0.7) +
  geom_smooth(color = "red") +
  labs(title = "Points per Game vs Salary", x = "Points per Game", y = "Salary")
```

## Summary Statistics

```{r}
nba |>
  group_by(Position) |>
  summarise(count = n(),
    mean_salary = mean(Salary, na.rm = TRUE),
    sd_salary = sd(Salary, na.rm = TRUE),
    min_salary = min(Salary, na.rm = TRUE),
    max_salary = max(Salary, na.rm = TRUE))
```

## Inferential Statistics

#### T-Test

```{r}
pg_vs_sg <- nba |>
  filter(Position %in% c("PG", "SG"))

t.test(Salary ~ Position, data = pg_vs_sg)
```

#### Linear Regression

```{r}
model <- lm(Salary ~ PTS, data = nba)
summary(model)
```

## Conclusion:

-   This analysis explored how NBA player performance relates to salary during the 2022–23 season. A histogram of points per game showed that most players score fewer than 20 points, with a few high scorers. A boxplot of salaries by position revealed that while all positions have outliers, there are differences in salary distributions depending on role. The scatterplot suggests a positive relationship between points per game and salary, which a linear regression model confirmed. The model found a statistically significant positive correlation: players who score more tend to earn higher salaries. Additionally, a t-test showed a significant salary difference between point guards and shooting guards, indicating that even within a single position group, roles can impact pay. Overall, both performance and specific position are associated with how much a player earns.
